Backward designing a reproducible framework for undergraduate data science research

### 2019-12-04 Rmd/discussion.Rmd
fixed headers
minor updates to discussion

## Discussion

### Benefits of our framework

1. build self-confidence, self-assessed proficiency in data analysis and statistics

@@

### How does the framework relate to ideas in Nolan and Temple Lang 2010??

### Critiques of our framework

@@

### Areas for improvement

@@

One shortcoming of our initial framework was the relative lack of emphasis on best practices for computational reproducibility. This is one area that we would like to rectify in future iterations. The University of Wisconsin-Madison has periodically offered a semester course in best practices for computationally reproducible research. We especially see version control systems, such as Git and Github, as essential tools for the modern data scientist.

>>>>

# Discussion

## Benefits of our framework

1. build self-confidence, proficiency in data analysis and statistics

@@

## How does the framework relate to ideas in Nolan and Temple Lang 2010??

## Critiques of our framework

@@

## Areas for improvement

@@

One shortcoming of our initial framework was the relative lack of emphasis on best practices for 
computational reproducibility. This is one area that we would like to rectify in 
future mentoring activities. The 
University of Wisconsin-Madison has periodically offered a semester course in best practices for 
computationally reproducible research (https://kbroman.org/Tools4RR/). We especially see collaborative 
version control systems, such as Git and 
Github, as essential tools for the modern data scientist.

1. assessment of data science skills

### 2019-12-04 Rmd/introduction.Rmd
upd
fixed intro
minor updates to discussion

- big data
- new areas of application of statistical methods - social media
- need for students to work with real data
- motivate statistical analysis and statistics research by real world scientific research questions (Box 1976)
- need for mentoring

@@

@nolan2010computing argue for students to work with real data. 
Working with real data allows them to develop skill not only in statistical analysis,
but also in data transfer from online sources, in data storage, and in using data 
from multiple file formats. In the case of Twitter data, tweets are stored in
Javascript Object Notation (JSON) [@tweet_json,@json].

>>>>

@@

@nolan2010computing argue for students to work with real data. 
Working with real data allows students to develop skill not only in statistical analysis,
but also in data transfer from online sources, in data storage, and in using data 
from multiple file formats. In the case of Twitter data, tweets are stored in
Javascript Object Notation (JSON) [@tweet_json,@json].

### 2019-12-04 Rmd/results.Rmd
fixed headers
updates
minor updates to discussion

## Results

We applied the project framework to our mentoring of two students. Both engaged in 12 months of mentored research during their senior year of undergraduate studies in statistics in 2015 and 2016.

### Student Outcomes

We subjectively assessed student outcomes through conversations in our weekly student research meetings.

One student later enrolled in a statistics graduate program, while the other pursued employment in health care analytics.

### Mentor Outcomes

### Scholarly outcomes & deliverables

>>>>

# Outcomes

We applied the project framework to our mentoring of two students. 
Both engaged in 12 months of mentored research during their senior year of 
undergraduate studies in statistics.

## Student Outcomes

We subjectively assessed student outcomes through conversations in our weekly 
student research meetings. Both students showed increases in confidence and ability to do 
data science research.

Both students secured positions in data science after graduation. 
One student later enrolled in a statistics graduate program, 
while the other pursued employment in health care analytics.

## Scholarly outcomes

Our scholarly contributions include the `parseTweetFiles` R package on Github
(https://github.com/rturn/parseTweetFiles) and presentations at conferences such as useR! 2016 and local poster sessions. Additionally, both students prepared end-of-project reports on their research.

We consider below
